Negative Selection in the 20th Century

### 2020-05-20 why-negative-selection.Rmd
Include all women in res_flb_weights. Fix bug which only used one PC in basic regressions; rewrite to use residualized variables. Interact scores with age_flb terciles.

mean_prop_wtd_unwtd <- res_weighted_all %>% 
      filter(sign(estimate_weighted) == sign(estimate_unweighted)) %>% 
      summarize(mean = mean(prop_wtd_unwtd)) %>% 
      pull(mean)

@@

We also weighted data by age at first live birth and age, among women only.
These weights are only basic attempts to correct for ascertainment bias. Effect
sizes might be increased further by more precise weighting.

@@

drake::loadd(res_flb_weights)
drake::loadd(res_sex)

bind_rows(
        Weighted   = res_flb_weights, 
        Unweighted = res_sex %>% filter(sex == "Female") %>% select(-subset, - sex),
        .id        = "Regression"
      ) %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(estimate, term, fill = Regression, 
        color = p.value < 0.05/66)) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2) +
        my_labs()
```

@@

Figure \@ref(fig:plot-age-flb) shows the results of *children* regressions for 
women only,
controlling for age at first live birth. Effect sizes are greatly reduced.
In `r opp_signed` out of `r length(score_names)` cases, they are of the 
opposite sign. The correlation between effect sizes
controlling for age at first live birth, and raw effect sizes, is `r cor_raw_flb`.

@@

>>>>

mean_prop_wtd_unwtd <- mean(res_weighted_all$prop_wtd_unwtd)

@@

```{r calc-res-flb-weights}

drake::loadd(res_flb_weights)
drake::loadd(res_sex)
res_flb_all <- left_join(
        res_flb_weights, 
        res_sex %>% filter(sex=="Female") %>% select(-subset, -sex), 
        by = "term", 
        suffix = c(".wt", ".nowt")
      ) %>% 
      mutate(prop_estimate = estimate.wt/estimate.nowt)

prop_wt_nowt <- mean(res_flb_all$prop_estimate)
pct_change_score <- function (score) {
  100 * (res_flb_all$prop_estimate[res_flb_all$term == score] - 1)
}
```

We also weighted data for women with children only, by age, education and age 
at first live birth. Weighting increased effect sizes on average by 
`r 100 * (prop_wt_nowt - 1)`%. Figure \@ref(fig:plot-flb-weights) shows the results. There 
are large changes for some variables, including EA3 
(`r pct_change_score("EA3_excl_23andMe_UK")`%), cognitive
ability (`r pct_change_score("cognitive_ability")`%) and
waist-hip ratio (WHR; `r pct_change_score("whr_combined")`%).

These weights are only basic attempts to correct for ascertainment bias. Effect
sizes might be increased further by more precise weighting.

@@

bind_rows(
        Weighted   = res_flb_weights, 
        Unweighted = res_flb_nowt %>% select(-subset),
        .id        = "Regression"
      ) %>% 
      mutate(
        term = pretty_names(term),
        term = fct_reorder(term, estimate, order_abs(1))
      ) %>% 
      ggplot(aes(estimate, term, fill = Regression, 
        color = p.value < 0.05/66)) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 2) +
        my_labs()
```

@@

Figure \@ref(fig:plot-age-flb) shows the results of *children* regressions for 
women only,
controlling for age at first live birth. Effect sizes are greatly reduced. Partly,
this is because the regressions exclude childless women. However,
in `r opp_signed` out of `r length(score_names)` cases, effect sizes actually
change sign with the controls. The correlation between effect sizes
controlling for age at first live birth, and raw effect sizes, is `r cor_raw_flb`.

Figure \@ref(fig:plot-age-flb-cross) shows *children* regressions estimated
separately for each tercile of age at first live birth. Several effects
are strikingly different. ADHD is selected for amongst the youngest
third of mothers, but selected against among the oldest two-thirds. Educational
attainment is selected for among the oldest two-thirds of mothers, but is not
significantly selected among the youngest third. Similarly, several PGS of
body measurements are selected against only among older mothers.

@@

```{r plot-age-flb-cross, fig.cap = "Effect sizes on number of children, by age at first live birth terciles"}

drake::loadd(res_age_flb_cross)

# this counts each cross term as a separate test:
n_regs <- as.double(nrow(res_age_flb_cross))

res_age_flb_cross %>% 
      mutate(
        score_name = pretty_names(score_name),
        score_name = fct_reorder(score_name, estimate, order_abs(1)),
        term       = gsub("age_flb_cat(.*):.*", "\\1", term)
      ) %>% 
      ggplot(aes(estimate, score_name, fill = term, 
        colour = p.value < 0.05/{{n_regs}})) +
        my_vline +
        pval_geom_point() +
        pval_scale() +
        my_fill_scale(n = 3, direction = -1, 
          guide = guide_legend(title = "Age at FLB")) +
        my_labs()
